---
title: "ST306_LA4_S_18_837"
author: "S/18/837"
date: "2023-09-26"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tinytex)
library(ggplot2)
```
```{r}
col_types <- "dccffi"
travel_data <- read_csv("../ST306_LA1_S_18_837/data/travel_data.csv"
                        , col_types = col_types)
```
```{r}
view(travel_data)
```
Main Question : 
What is the most common mode of transportation used to travel from home to work?
```{r}
x1 <- travel_data%>%
  group_by(mode_of_trans)%>%
  summarise(Count = sum(count))%>%
  select(mode_of_trans, Count)%>%
  arrange(desc(Count))
x1
```

```{r}
x1 %>%
  ggplot(aes(x=mode_of_trans, y=Count, fill = mode_of_trans))+geom_bar(stat = "identity")+
  labs(title = "Distribution of Mode of Transportation",x = "Mode of Transportation",
       y = "Count")
```
Conclusion :	The most common mode of transportation in the travel_data data set is "drove_alone".



Sub Question 1:
Q1)What is the most common mode of transportation used to travel from home to
work in 2018?
```{r}
Q1 <- travel_data %>%
  filter(year == 2018) %>%
  group_by(mode_of_trans) %>%
  summarise(Count = sum(count)) %>%
  select(mode_of_trans, Count)
Q1
```
```{r}
Q1 %>% 
  ggplot(aes(x=mode_of_trans, y=Count,fill = mode_of_trans))+geom_bar(stat = "identity") +
  labs(title = "Distribution of Mode of Transportation in 2018",x = "Mode of Transportation",
       y = "Count")
```
Conclusion : The most common mode of transportation in 2018 in the travel_data data set is "drove_alone".



Sub Question 2:
Q2)  How does the mode of transportation vary by region?

```{r}
travel_data %>%
  group_by(region, mode_of_trans) %>%
  summarise(count = sum(count)) %>%
  ggplot(aes(x = region, y = count, fill = mode_of_trans)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Mode of Transportation Distribution by Region",
       x = "Region",
       y = "Count") 
```
Conclusion : In all regions "drove alone" is the more commonly used method to travel and "bicycle" has the lowest                 rate.               


Sub Question 3: 
Q3)How does the mode of transportation vary by gender?
```{r}
travel_data%>%
  group_by(gender, mode_of_trans) %>%
  summarise(count = sum(count)) %>%
  ggplot(aes(x = gender, y = count, fill = mode_of_trans)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Mode of Transportation Distribution by Gender",
       x = "Gender",
       y = "Count")
```
Conclusion : Generally, "drove_alone" is the most common mode for both genders, but the proportions differ.


Sub Question 4:
Q4) What is the most common mode of transport in Michigan in 2018?
```{r}
travel_data %>%
  filter(year == 2018 & state == "Michigan")%>%
  group_by(mode_of_trans)%>%
  summarise(Count = sum(count))
```

```{r}
travel_data %>%
  filter(year == 2018 & state == "Michigan")%>%
  group_by(mode_of_trans)%>%
  summarise(Count = sum(count))%>%
  ggplot(aes(x=mode_of_trans, y=Count, fill = mode_of_trans))+geom_bar(stat = "identity") +
    labs(title = "Distribution of Mode of Transportation of Michigan in 2018",x = "Mode of Transportation",
       y = "Count")
```

Conclusion : The more commonly transportation mode in Michigan in 2018 is drove alone followed by carpooled, walking,
            public transportation, taxi cab or other methods and bicycle.
            
            
Sub Question 5: 
Q5) Are there any trends in transportation mode usage over the years?
```{r}
travel_data %>%
  group_by(year, mode_of_trans) %>%
  summarise(total_count = sum(count)) %>%
  ggplot(aes(x = year, y = total_count, color = mode_of_trans)) +
  geom_line() +
  labs(title = "Trends in Transportation Mode Usage Over the Years",
       x = "Year",
       y = "Total Count")
```
Conclusion : We can see a significant increment of driving alone and carpooled methods but others more likely to the             same. 
